\documentclass{article}


\title{One reason to version control source files only}

\begin{document}

\maketitle

This is the first sentence.
This is the 2 th sentence.
This is the 3 th sentence.
This is the 4 th sentence.
This is the 5 th sentence.
This is the 6 th sentence.
This is the 7 th sentence.
This is the 8 th sentence.
This is the 9 th sentence.
This is the 10 th sentence.
This is the 11 th sentence.
This is the 12 th sentence.
This is the 13 th sentence.
This is the 14 th sentence.
This is the 15 th sentence.
This is the 16 th sentence.
This is the 17 th sentence.
This is the 18 th sentence.
This is the 19 th sentence.
This is the 20 th sentence.
This is the 21 th sentence.
This is the 22 th sentence.
This is the 23 th sentence.
This is the 24 th sentence.
This is the 25 th sentence.
This is the 26 th sentence.
This is the 27 th sentence.
This is the 28 th sentence.
This is the 29 th sentence.
This is the 30 th sentence.
This is the 31 th sentence.
This is the 32 th sentence.
This is the 33 th sentence.
This is the 34 th sentence.
This is the 35 th sentence.
This is the 36 th sentence.
This is the 37 th sentence.
This is the 38 th sentence.
This is the 39 th sentence.
This is the 40 th sentence.
This is the 41 th sentence.
This is the 42 th sentence.
This is the 43 th sentence.
This is the 44 th sentence.
This is the 45 th sentence.
This is the 46 th sentence.
This is the 47 th sentence.
This is the 48 th sentence.
This is the 49 th sentence.
This is the 50 th sentence.
This is the 51 th sentence.
This is the 52 th sentence.
This is the 53 th sentence.
This is the 54 th sentence.
This is the 55 th sentence.
This is the 56 th sentence.
This is the 57 th sentence.
This is the 58 th sentence.
This is the 59 th sentence.
This is the 60 th sentence.
This is the 61 th sentence.
This is the 62 th sentence.
This is the 63 th sentence.
This is the 64 th sentence.
This is the 65 th sentence.
This is the 66 th sentence.
This is the 67 th sentence.
This is the 68 th sentence.
This is the 69 th sentence.
This is the 70 th sentence.
This is the 71 th sentence.
This is the 72 th sentence.
This is the 73 th sentence.
This is the 74 th sentence.
This is the 75 th sentence.
This is the 76 th sentence.
This is the 77 th sentence.
This is the 78 th sentence.
This is the 79 th sentence.
This is the 80 th sentence.
This is the 81 th sentence.
This is the 82 th sentence.
This is the 83 th sentence.
This is the 84 th sentence.
This is the 85 th sentence.
This is the 86 th sentence.
This is the 87 th sentence.
This is the 88 th sentence.
This is the 89 th sentence.
This is the 90 th sentence.
This is the 91 th sentence.
This is the 92 th sentence.
This is the 93 th sentence.
This is the 94 th sentence.
This is the 95 th sentence.
This is the 96 th sentence.
This is the 97 th sentence.
This is the 98 th sentence.
This is the 99 th sentence.
This is the 100 th sentence.

\end{document}